GEEPERs: Principal Stratification using Principal Scores and Stacked Estimating Equations

### 2022-12-13 writeUps/paper.tex

In each run of the simulation, we simulated three indpendent standard normal covariates, $x_k$ $k=1,\dots,3$. However, only the first two covariates were ``observed,'' i.e. were included in the analysis model.
Given the covariates, the (true) principal scores were defined as
\begin{equation*}
  \ppi=logit^{-1}\left[\alpha(x_{1i}-x_{2i}+x_{3i})\right]
\end{equation*}
where $\alpha$ is a manipulated factor.
When $\alpha$ was higher, $\st$ was more easily predicted by covariates; when $\alpha=0$, Assumption \ref{ass:vps}, was violated, i.e. $Var(\pp)=0$.
Then principal stratum $\sti\in \{0,1\}$ was simulated as $S_i\sim Bern(\ppi)$

@@

\textbf{\emph{Interactions between $\st$ or $Z$ and covariates:}} The last two factors---the presence of interactions between covariates and $\st$ and $Z$, respectively---test the performance of a model that does not include interactions when that modeling assumption is false.
When neither interaction was present, $\gamma_2=\gamma_3=0$, then $\gamma_1=1/\sqrt{6}$.
Then, since  $Var(\gamma_1(x_1+x2+x_3))=1/2$ and $Var(\epsilon_i)=1/2$, covariates explained half of the variance of $Y$ within treatment group and principal stratum, and observed covariates $x_1$ and $x_2$ explained 1/3 of that variance.
When there was an interaction between covariates and $\sti$, then $\gamma_1=3/(4\sqrt{6})$ and $\gamma_2=1/(2\sqrt{6})$, so that the slopes for $x_1$ and $x_2$ varied by half their magnitude between principal strata, and on average they were equal to $1/\sqrt{6}$, as in the no-interaction case.
Interactions between covariates and $Z$ were controlled by $\gamma_3$: when there were interactions, $\gamma_3=1/(2sqrt{6})$, otherwise $\gamma_3=0$.

@@

\begin{figure}[!ht]
  \centering
  \includegraphics[width=\textwidth]{../simFigs/boxplots.pdf}
  \caption{Violin and scatterplots of 500 simulation estimation errors for estimators described in Section~\ref{sec:simMods} under varying conditions. For all plots, $n=500$; there were 500 replications for each set of conditions. Annotations at the upper and lower edges of the plot indicate the number of extreme outliers excluded from the plotting area.}
  \label{fig:boxplots}
\end{figure}

Figure \ref{fig:boxplots} shows violin plots, layered on top of jittered scatterplots, showing estimation error for GEEPERs, the mixture estimator, and the PSW estimator of $\eff1$ under various sets of conditions, and Table \ref{tab:coverage} shows empirical covarage of nominal 95\% confidence interval estimates under a somewhat wider set of circumstances when $\muc1=0.3$.
The leftmost panel of Figure \ref{fig:boxplots} and the top line of Table \ref{tab:coverage} show results conditions favorable to both the mixture model and to GEEPERs---normal residuals and no interactions in the data generating model between covariates and either $S$ or $Z$.
In this case, GEEPERs is roughly unbiased, but less precise than the other two methods---when $\alpha=0.2$ it is much noisier, and when $\alpha=0.5$ it is only slightly less precise.
The mixture estimator's bias depends on $\muc1$---when $\muc1=0$, the bias is slight, and when $\muc1=0.3$ the bias is more substantial. This is somewhat surprising, because when $\muc1=0$, there is no separation between the principal strata in the control group, and previous literature \citep{griffin2008application} has identified this scenario as a particular challenge for mixture modeling.
The PSW estimator is also biased---due, presumably, to the violation of Principal Ignorability, Assumption \ref{ass:PI}---inherent in the data generating model. On the other hand, the PSW estimator has easily the lowest sampling variance of the three estimators.
The middle panel of Figure \ref{fig:boxplots} and the fifth line of Table \ref{tab:coverage} show a similar set of results for when the residuals in \eqref{eq:y-sim} are uniformly distributed.
In this case, GEEPERs and PSW estimator behave in a similiar fashion to the normal case, but the mixture estimator shows a bimodal pattern--sometimes over-estimating $\eff1$ and sometimes underestimating, but rarely estimating $\eff1$ accurately.

\sloppy
The rightmost panel of Figure \ref{fig:boxplots} returns to the case of normal residuals---and only shows results when $\muc1=0.3$ and $\alpha=0.5$---but introduces interactions in the data generating model between covariates and either $\st$, $Z$, or both. That is, the coefficients on the covariates in the outcome model vary between principal strata and/or between treatment groups.
Presence of interactions could undermine the GEEPERs by inducing a relationship between principal scores and potential outcomes, even after adjusting the outcomes for covariate effects.
Fortunately, the presence of interactions in the data generating model in this scenario appear to make little difference, though RMSEs were slightly higher when interactions between $Z$ and covariates are present.
Table \ref{tab:rmse} shows that when $\alpha=.2$, GEEPERs RMSE can be substantially higher when $Z$ interactions are present than when they are not; plots in the online appendix suggest that this is due to an increase in sampling variance and kurtosis rather than bias.

Table \ref{tab:coverage} shows estimates of the empirical coverage of nominal 95\% confidence intervals for GEEPERs and the mixture estimator of $\eff1$, to evaluate their respective standard error estimates and the associated statistical inference.
When $\alpha=0.5$, GEEPERs confidence intervals consistently achieve their nominal levels\footnote{With 500 replications and true coverage probabilities of 0.95, the standard error for empirical coverage rates is approximately 1 percentage point.}, whereas confidence intervals from mixture estimators undercover in the presence of interactions between $Z$ and $X$, or when residuals are non-normal.
When $\alpha=0$ or 0.2, GEEPERs confidence intervals undercover in the presence of interactions between $X$ and $Z$.
Across all conditions in the table, the GEEPERs confidence intervals have higher empirical coverage than their mixture-model counterparts.

@@

\multirow{-4}{*}{\raggedright\arraybackslash Uniform} & Yes & Yes & 0.83 & 0.15 & 0.88 & 0.21 & 0.94 & 0.53\\
\bottomrule

@@

In a recent educational field experiment, middle school students completing their math schoolwork on a computer program were randomized between four conditions. In all four conditions, students worked on their assigned math program in class on 12 different occasions---a prior assessment, a mid-test, a post-test, and nine learning sessions---spaced throughout the school year.  
The four conditions ranged from a game---DragonBox---to the active control condition in which students simply used the computer to work on a sequence of math problems. 
In the DragonBox game, students learned to solve algebraic equations by isolating a box containing a dragon; as students progressed in the game, pictures of monsters were replaced by mathematical symbols. 
Another condition, the From Here to There (FH2T) program, used visual cues such as color and spacing to teach algebraic concepts, and allowed students to solve equations by manipulating expressions via dragging and dropping.
In the ASSISTments condition, students used computers to work through a series of math problems taken from the district's textbook. While working, students had access to a series hints for each problem, culminating in a ``bottom out'' hint that contained the answer. Students in the ASSISTments condition were given immediate feedback on their answers and had to enter the correct answer in order to proceed to the next problem. 
In the active control condition, labeled ``business as usual'' (BAU), the computer work mimicked traditional pencil-and-paper mathematics coursework---students worked through a series of math problems without access to hints and received feedback only after completing the assignment. 
In this illustration, we will compare the ASSISTments condition (which we arbitrarily label ``treatment'')---the only condition featuring bottom-out hints---with each of the three others separately. Our goal is to estimate different treatment effects for groups of students who would, if assigned to the ASSISTments condition, request few or many bottom out hints.

The literature on hints in online learning is mixed \citep[see, e.g.,][]{aleven2016help,goldin2012learner,sales2021student} in particular when it comes to requesting a bottom-out hint (called ``bottoming out''). On the one hand, hints that include the correct answer are essentially worked examples, which can be beneficial to learning \citep[e.g.]{sweller1985use}. On the other hand, they allow students to ``game the system,'' by simply requesting enough hints to see the right answer, and use the software without ever working to solve any problems \citep[e.g.][]{guo2008trying}. However, in this experiment, the ability to bottom out was only one of several differences between the treatment conditions. One way to disentangle the role of bottoming out in any treatment effect would be to estimate average treatment effects separately for subjects who would (if assigned to treatment) bottom out, and for subjects who would not. If hints (that do not supply the correct answer) and error messages are helpful, but the ability to bottom out is harmful to learning, then we might expect to estimate a positive effect of being assigned to the feedback condition for students who do not bottom out, but a negative effect for students who do. If hints are helpful for learning only if they can be read as worked examples, then we might expect no effect of condition on students who do not bottom out, but a positive effect for students who do. (These hypothetical results may be masked by other treatment effect heterogeneity between students who would or wouldn't bottom out, but the results could still be informative.)

For the purpose of demonstrating GEEPERs, we gathered dichotomized data on %which students in the treatment condition requested at least one bottom-out hint.
whether students in the treatment condition requested more than the median number of bottom-out hints (11).
That is, $\st=1$ for students who, if assigned to the treatment condition, would request at more than 11 bottom-out hints (``bottom-outers''), and $\st=0$ for students who would request 11 or fewer (``non-bottom-outers''). %least one bottom-out hint, and $\st=0$ for students who would not.
The outcome of interest $Y$ in our example is students' total scores on a ten-item posttest completed within the online tutoring system; $Y$ is an integer ranging from 0 to 10.
The goal of our analysis will be to determine the principal effects $\tau_1=\EE[\yt-\yc|\st=0]$ and $\tau_0=\EE[\yt-\yc|\st=0]$, the average effects of assignment to the treatment condition for bottom-outers and non-bottom-outers.

We also had access to data on a number of baseline covariates: total scores on a 10-item pretest, fifth-grade standardized test scores, race/ethnicity (white, Asian, Hispanic/Latino, or Other), indicator variables for if students have an early intervention program (EIP) or individualized educational plan (IEP) in place, if students are learning English as a second language (ESOL), students' average time on task for the pretest (logged), total scores from baseline evaluations of students' math anxiety, negative reaction to math, and numerical confidence, and an indicator for weather students began the school year in remote or in-person instruction.
Students with missing pretest scores were dropped from the analysis; missing data in other covariates was imputed with a Random Forest algorithm \citep{missForest}.
%Table \ref{tab:tab1} describes each covariate--as well as the sucess of the randomForest imputation algorithm.
%% \begin{comment}
%% \begin{table}
%%   \centering
%%   \input{table1}
%%   \caption{Descriptive statistics on baseline covariates, $S$, and $Y$ for the analysis data. Values are mean (SD) unless otherwise noted. Imputation error is the ``out of box'' error estimate from the Random Forest imputation algorithm---root mean squared error divided by the observed standard deviation for continuous variables and misclassification proportion for categorical variables.}
%%   \label{tab:tab1}
%% \end{table}
%% \end{comment}
%% u

@@

\subsection{Principal Score Model}\label{sec:psMod}

\begin{figure}
  \centering
  \includegraphics{../figure/psModCoef.jpg}
  \caption{Principal Score Model Coefficients}
  \label{fig:psMod}
\end{figure}

We considered several models for the principal score, and full results (which are largely similar to each other) can be found in the online appendix. Here we present results from our model of choice, a logistic regression of observed $S$ on baseline variables chosen with a combination of expert judgement and backwards selection based on the AIC. These included school fixed effects and a set of student-level demographic and educational data: fifth-grade mathematics standardized test scores, %Gender, Race/Ethnicity,
indicator variables for special education status (IEP), whether the student was taking English as a second language (ESOL), whether the student was classified as gifted, and whether students attended a full year of school in 5th grade. Finally, we also included data collected from students at the beginning of the study: the natural logarithm of the  time spent on task during the pretest, scores on baseline measurements of mathematics anxiety and self-efficacy and perceptual sensitivity, along with a missingness indicator for these measurements.
Binned residual plots \citep{gelman2006data} from an initial model fit indicated a curvilinear relationship between perceptual sensitivity measurements (modeled as linear) and model errors, so in the final model we included perceptual sensitivity as a quadratic polynomial.
The model was fit to data from students in the treatment group, for whom bottom-out status ($S$) was observed; hence, the same model fit was used for all three treatment contrasts.

The model was fairly successful in distinguishing bottom-outers from non-bottom-outers---the AUC, evaluated using out-of-sample predictions in a 10-fold cross-validation, was roughly 0.795. That is, in roughly 80\% of bottom-outer/non-bottom-outer pairs of subjects, the bottom-outer will have a higher predicted probability from the model.
Table \ref{fig:psMod} show the model's coefficient estimates. Notably, conditional on the other covariates, students with higher fifth-grade standardized test scores were less likely to request many bottom out hints, as were students who spent little time on task during the pretest, while IEP and ESOL students, and students without baseline measurement data, were relatively unlikely to be bottom-outers. The relationship between perceptual sensitivity and bottom out requests was not monotonic---students with average levels were more likely than those with more extreme high or low levels to be bottom-outers; bottom-outers tended to have lower math self-efficacy, but the data are also consistent with a small positive relationship, and there appears to be little to no relationship between math anxiety and bottom-outer status at all.

\subsection{Outcome Regression and Principal Effects}\label{sec:outMod}
\begin{figure}
  \centering
  \includegraphics{../figure/prinEffs.pdf}
  \caption{Principal Effects}
  \label{fig:effects}
\end{figure}

Figure \ref{fig:effects} plots the estimated principal effects for bottom-outers and non-bottom-outers, using two different models---alongside estimates of the overall average treatment effect, labeled ``ATE.'' In both models, effects were adjusted for all of the variables that were in the principal score model, in addition to students' pretest scores; also, the models substituted class fixed effects for school fixed effects.\footnote{Why use a coarser set of fixed effects for the principal score model than the outcome regression? There were two reasons: first, the principal score model was fit to a smaller sample (only students in the ``Instant'' condition), so some classrooms only included one student, which can lead to poor fit. Second, the asymptotics of logistic regression are more restricted than OLS, and so logistic regression tends to perform poorly when the number of parameters increases with the sample size \citep{agresti}.}
The two sets of estimated differed in that the first assumed that coefficients on covariates were equal across principal strata, whereas the second set of results allowed the coefficients on pretest and 5th-grade standardized test scores---the most important predictors of outcomes and principal strata membership, respectively---to vary between strata.

The results suggest little difference, if any, between the effects in the two principal strata.
Estimated effects compared to BAU or DragonBox were slightly lower for bottom-outers than for non-bottom-outers, and estimated effects compared to FH2T were higher for non-bottom-outers.
However, none of these differences is statistically significant, so opposite patterns, or no difference between principal effects at all, would also be consistent with the data.
In fact, 95\% confidence intervals for all principal effects include both positive and negative principal effects.
There is no discernible difference between estimated principal effects when coefficients on pretest and grade-5 standardized tests were or were not allowed to vary between strata, though standard errors from the models including an interaction were somewhat higher.

\subsection{Mixture Model and Principal Score Weighting}
\begin{figure}
  \centering
  \includegraphics{../figure/compareMethods.pdf}
  \caption{Principal Effects}
  \label{fig:compare}
\end{figure}

Figure \ref{fig:compare} shows principal effect estimates from GEEPERs (without the $S$-interactions) compared to analogous estimates using Bayesian mixture modeling and PSW.
All three sets of estimates relied on the same principal score model described above, in Section \ref{sec:psMod}, though in the mixture model it was fit simultaneously with effect estimation using a Markov Chain Monte Carlo algorithm \citep{rstan}.
The Bayesian mixture model assumed that outcomes were normally distributed, conditional on covariates (the same as in Section \ref{sec:outMod}) and principal stratum, with a residual standard deviation that varied between principal strata. This modeling assumption is necessarily false, since post-test scores were integers between 0 and 10. All model parameters with support in $\mathbb{R}$ were given standard normal priors, while standard deviation parameters were given half-standard normal priors.
The PSW estimators did not include any covariate outcome modeling. Standard errors for the PSW estimates were estimated using the bootstrap, with resampling done within schools.
For all three sets of estimators, approximate 95\% confidence intervals were estimated by subtracting twice the standard error from the point estimate (or twice the posterior standard deviation from the posterior mean, in the Bayesian mixture model).

In this case, all three estimation strategies largely agreed. This may suggest that the estimators' various assumptions---additivity of treatment assignment effect for GEEPERs and the mixture model, normal residuals for the mixture model, and principal ignorability for PSW---were all innocuous and may have held approximately. Alternatively, violations of the different assumptions may have led to similar errors.

>>>>

In each run of the simulation, we simulated three independent standard normal covariates, $x_k$ $k=1,\dots,3$. However, only the first two covariates were ``observed,'' i.e. were included in the analysis model.
Given the covariates, the (true) principal scores were defined as
\begin{equation*}
  \ppi=logit^{-1}\left[\alpha(x_{1i}-x_{2i}+x_{3i})\right]
\end{equation*}
where $\alpha$ is a manipulated factor.
When $\alpha$ was higher, $\st$ was more easily predicted by covariates; when $\alpha=0$, Assumption \ref{ass:vps}, was violated, i.e. $Var(\pp)=0$.
Then principal stratum $\sti\in \{0,1\}$ was simulated as $S_i\sim Bern(\ppi)$

@@

\textbf{\emph{Interactions between $\st$ or $Z$ and covariates:}} The last two factors---the presence of interactions between covariates and $\st$ and $Z$, respectively---test the performance of a model that does not include interactions when that modeling assumption is false.
When neither interaction was present, $\gamma_2=\gamma_3=0$, then $\gamma_1=1/\sqrt{6}$.
Then, since  $Var(\gamma_1(x_1+x2+x_3))=1/2$ and $Var(\epsilon_i)=1/2$, covariates explained half of the variance of $Y$ within treatment group and principal stratum, and observed covariates $x_1$ and $x_2$ explained 1/3 of that variance.
When there was an interaction between covariates and $\sti$, then $\gamma_1=3/(4\sqrt{6})$ and $\gamma_2=1/(2\sqrt{6})$, so that the slopes for $x_1$ and $x_2$ varied by half their magnitude between principal strata, and on average they were equal to $1/\sqrt{6}$, as in the no-interaction case.
Interactions between covariates and $Z$ were controlled by $\gamma_3$: when there were interactions, $\gamma_3=1/(2sqrt{6})$, otherwise $\gamma_3=0$.
Either type of interaction violates Assumption \ref{ass:rci} since if $\gamma_2\ne 0$ or $\gamma_3 \ne 0$, there is no single vector $\bm{\gamma}$ that can fully capture the dependence of $\yc$ and $\yt$ on $S$---instead, the dependence varies with $\st$ and/or with $Z$.

@@

\begin{figure}[!ht]
  \centering
  \includegraphics[width=\textwidth,clip]{../simFigs/boxplots.pdf}
  \caption{Violin and scatterplots of 500 simulation estimation errors for estimators described in Section~\ref{sec:simMods} under varying conditions. For all plots, $n=500$; there were 500 replications for each set of conditions. Annotations indicate the number of extreme outliers excluded from the plotting area.}
  \label{fig:boxplots}
\end{figure}

Figure \ref{fig:boxplots} shows violin plots, layered on top of jittered scatterplots, showing estimation error for GEEPERs, the mixture estimator, and the PSW estimator of $\eff1$ under various sets of conditions, and Table \ref{tab:coverage} shows empirical covarage of nominal 95\% confidence interval estimates under a somewhat wider set of circumstances when $\muc1=0.3$.
The leftmost panel of Figure \ref{fig:boxplots} and the top line of Table \ref{tab:coverage} show results conditions favorable to both the mixture model and to GEEPERs---normal residuals and no interactions in the data generating model between covariates and either $S$ or $Z$.
In this case, GEEPERs is roughly unbiased, but less precise than the other two methods---when $\alpha=0.2$ it is much noisier, and when $\alpha=0.5$ it is only slightly less precise.
The mixture estimator's bias depends on $\muc1$---when $\muc1=0$, the bias is slight, and when $\muc1=0.3$ the bias is more substantial. This is somewhat surprising, because when $\muc1=0$, there is no separation between the principal strata in the control group, and previous literature \citep{griffin2008application} has identified this scenario as a particular challenge for mixture modeling.
The PSW estimator is also biased---due, presumably, to the violation of Principal Ignorability, Assumption \ref{ass:PI}---inherent in the data generating model. On the other hand, the PSW estimator has easily the lowest sampling variance of the three estimators.
The middle panel of Figure \ref{fig:boxplots} and the fifth line of Table \ref{tab:coverage} show a similar set of results for when the residuals in \eqref{eq:y-sim} are uniformly distributed.
In this case, GEEPERs and PSW estimator behave in a similar fashion to the normal case, but the mixture estimator shows a bimodal pattern--sometimes over-estimating $\eff1$ and sometimes underestimating, but rarely estimating $\eff1$ accurately.
Credible intervals from the mixture model exhibit severe undercoverage, while GEEPERs confidence intervals achieve at least their nominal level.

%\sloppy
The rightmost panel of Figure \ref{fig:boxplots} returns to the case of normal residuals---and only shows results when $\muc1=0.3$ and $\alpha=0.5$---but introduces interactions in the data generating model between covariates and either $\st$, $Z$, or both. That is, the coefficients on the covariates in the outcome model vary between principal strata and/or between treatment groups, violating Assumption \ref{ass:rci}..
%Presence of interactions could undermine the GEEPERs by inducing a relationship between principal scores and potential outcomes, even after adjusting the outcomes for covariate effects.
Fortunately, the presence of interactions in the data generating model in this scenario appear to make little difference when $\alpha=0.5$; however, Table \ref{tab:coverage} shows that when $\alpha=0$ or 0.2, and there are interactions between $\bx$ and $Z$ (i.e. $\gamma_3\ne 0$) GEEPERs 95\% confidence intervals tend to under-cover. Nevertheless, coverage of GEEPERs intervals is nearly always substantially higher than coverage of corresponding 95\% credible intervals from mixture models.%\footnote{}

@@

\multirow{-4}{*}{\raggedright\arraybackslash Uniform} & Yes & Yes & 0.83 & 0.15 & 0.88 & 0.21 & 0.94 & 0.53\\
\bottomrule
\multicolumn{9}{p{0.9\textwidth}}{\footnotesize Standard error $\approx 1$ percentage point, based on 500 replications and true coverage probabilities of 0.95.}

@@

In a recent educational field experiment, middle school students completing their math schoolwork on a computer program were randomized between four conditions. In all four conditions, students worked on their assigned math program in class on 12 different occasions---a prior assessment, a mid-test, a post-test, and nine learning sessions---spaced throughout the school year.  
The four conditions included two gamified programs---DragonBox and From Here to There (FH2T)---and two programs in which students used computers to work through a series of algebra problems taken from the district's textbook---``Assistments'' and an active control called ``Business as Usual'' (BAU).  % the active control condition in which students simply used the computer to work on a sequence of math problems. 
%In the DragonBox game, students learned to solve algebraic equations by isolating a box containing a dragon; as students progressed in the game, pictures of monsters were replaced by mathematical symbols. 
%Another condition, the From Here to There (FH2T) program, used visual cues such as color and spacing to teach algebraic concepts, and allowed students to solve equations by manipulating expressions via dragging and dropping.
%In the ASSISTments condition, students used computers to work through a series of math problems taken from the district's textbook. 
Students in the ASSISTments condition had access to a series hints for each problem, culminating in a ``bottom out'' hint that contained the answer and were given immediate feedback on their answers. % and had to enter the correct answer in order to proceed to the next problem. 
In the BAU condition students had no access to hints and received feedback only after completing the assignment. 
In this illustration, we will compare the ASSISTments condition (which we arbitrarily label ``treatment'')---the only condition featuring bottom-out hints---with each of the three others separately. Our goal is to estimate different treatment effects for groups of students who would, if assigned to the ASSISTments condition, request few or many bottom out hints.

The literature on hints in online learning is mixed \citep[see, e.g.,][]{aleven2016help,goldin2012learner,sales2021student} in particular when it comes to requesting a bottom-out hint (called ``bottoming out''). On the one hand, hints that include the correct answer are essentially worked examples, which can be beneficial to learning \citep[e.g.]{sweller1985use}. On the other hand, they allow students to ``game the system,'' by simply requesting enough hints to see the right answer, and use the software without ever working to solve any problems \citep[e.g.][]{guo2008trying}. However, in this experiment, the ability to bottom out was only one of several differences between the treatment conditions. One way to disentangle the role of bottoming out in any treatment effect would be to estimate average treatment effects of being assigned to the ASSISTments condition, relative to any of the other three, separately for subjects who would (if assigned to treatment) bottom out frequently, and for subjects who would not. 
If the availability of bottom-out hints plays an important positive role in ASSISTments's efficacy, we might expect larger effects of condition on students who would bottom out frequently than for those who would not; if bottom-out hints are harmful, we might expect the opposite. 
%If hints (that do not supply the correct answer) and error messages are helpful, but the ability to bottom out is harmful to learning, then we might expect to estimate a positive effect of being assigned to the feedback condition for students who do not bottom out, but a negative effect for students who do. If hints are helpful for learning only if they can be read as worked examples, then we might expect no effect of condition on students who generally do not bottom out, but a positive effect for students who do. 
(These hypothetical results may be masked by other treatment effect heterogeneity between students who would or wouldn't bottom out, but the results could still be informative.)

\sloppy
For the purpose of demonstrating GEEPERs, we gathered dichotomized data on %which students in the treatment condition requested at least one bottom-out hint.
whether students in the treatment condition requested more than the median number of bottom-out hints (11).
That is, $\st=1$ for students who, if assigned to the treatment condition, would request at more than 11 bottom-out hints (``bottom-outers''), and $\st=0$ for students who would request 11 or fewer (``non-bottom-outers''). %least one bottom-out hint, and $\st=0$ for students who would not.
The outcome of interest $Y$ in our example is students' total scores on a ten-item posttest completed within the online tutoring system; $Y$ is an integer ranging from 0 to 10.
The goal of our analysis will be to determine the principal effects $\tau_1=\EE[\yt-\yc|\st=0]$ and $\tau_0=\EE[\yt-\yc|\st=0]$, the average effects of assignment to the treatment condition for bottom-outers and non-bottom-outers.

We also had access to data on a number of baseline covariates, including prior achievement, demographics, baseline measurements of student attitudes towards math and an indicator for whether students began the school year in remote or in-person instruction.
Students with missing pretest scores were dropped from the analysis; missing data in other covariates was imputed with a Random Forest algorithm \citep{missForest}.

@@

Due to privacy regulations, the data are not publicly available; however, interested analysts may access anonymized data by emailing the study's principal investigator at [REDACTED] and signing a privacy agreement.
Replication code for the analysis is available at \url{www.github.com/[Redacted]}.

\subsection{Data Analysis}
\subsubsection{Principal Score Model}\label{sec:psMod}

% \begin{figure}
%   \centering
%   \includegraphics{../figure/psModCoef.jpg}
%   \caption{Principal Score Model Coefficients}
%   \label{fig:psMod}
% \end{figure}

We estimated principal scores using logistic regression of observed $S$ on schol fixed effects and a set of baseline variables chosen with a combination of expert judgement and backwards selection based on the AIC, and evaluated model fit with binned residual plots \citep{gelman2006data}.
The model was fit to data from students in the treatment group, for whom bottom-out status ($S$) was observed; hence, the same model fit was used for all three treatment contrasts.
The model was fairly successful in distinguishing bottom-outers from non-bottom-outers---the AUC, evaluated using out-of-sample predictions in a 10-fold cross-validation, was roughly 0.795. That is, in roughly 80\% of bottom-outer/non-bottom-outer pairs of subjects, the bottom-outer will have a higher predicted probability from the model.

% \subsubsection{Outcome Regression and Principal Effects}\label{sec:outMod}
% \begin{figure}
%   \centering
%   \includegraphics{../figure/prinEffs.pdf}
%   \caption{Principal Effects}
%   \label{fig:effects}
% \end{figure}

% Figure \ref{fig:effects} plots the estimated principal effects for bottom-outers and non-bottom-outers, using two different models---alongside estimates of the overall average treatment effect, labeled ``ATE.'' In both models, effects were adjusted for all of the variables that were in the principal score model, in addition to students' pretest scores; also, the models substituted class fixed effects for school fixed effects. %\footnote{Why use a coarser set of fixed effects for the principal score model than the outcome regression? There were two reasons: first, the principal score model was fit to a smaller sample (only students in the ``Instant'' condition), so some classrooms only included one student, which can lead to poor fit. Second, the asymptotics of logistic regression are more restricted than OLS, and so logistic regression tends to perform poorly when the number of parameters increases with the sample size \citep{agresti}.}
% The two sets of estimated differed in that the first assumed that coefficients on covariates were equal across principal strata, whereas the second set of results allowed the coefficients on pretest and 5th-grade standardized test scores---the most important predictors of outcomes and principal strata membership, respectively---to vary between strata.

% The results suggest little difference, if any, between the effects in the two principal strata.
% Estimated effects compared to BAU or DragonBox were slightly lower for bottom-outers than for non-bottom-outers, and estimated effects compared to FH2T were higher for non-bottom-outers.
% However, none of these differences is statistically significant, so opposite patterns, or no difference between principal effects at all, would also be consistent with the data.
% In fact, 95\% confidence intervals for all principal effects include both positive and negative principal effects.
% There is no discernible difference between estimated principal effects when coefficients on pretest and grade-5 standardized tests were or were not allowed to vary between strata, though standard errors from the models including an interaction were somewhat higher.

\begin{figure}
  \centering
  \includegraphics{../figure/compareMethods.pdf}
  \caption{Principal Effects}
  \label{fig:compare}
\end{figure}

Figure \ref{fig:compare} shows principal effect estimates from GEEPERs compared to analogous estimates using Bayesian mixture modeling and PSW, alongside estimates of the overall average treatment effects, labeled ``ATE.''.
All three sets of estimates relied on the same principal score model described above, in Section \ref{sec:psMod}, though in the mixture model it was fit simultaneously with effect estimation using a Markov Chain Monte Carlo algorithm \citep{rstan}.
Outcome regressions in GEEPERs and the mixture model included all of the covariates that were in the principal score model, in addition to students' pretest scores; also, the models substituted class fixed effects for school fixed effects.

The Bayesian mixture model assumed that outcomes were normally distributed, conditional on covariates (the same as in Section \ref{sec:outMod}) and principal stratum, with a residual standard deviation that varied between principal strata. This modeling assumption is necessarily false, since post-test scores were integers between 0 and 10. All model parameters with support in $\mathbb{R}$ were given standard normal priors, while standard deviation parameters were given half-standard normal priors.

The PSW estimators did not include any covariate outcome modeling. Standard errors for the PSW estimates were estimated using the bootstrap, with resampling done within schools.

For all three sets of estimators, approximate 95\% confidence intervals were estimated by adding and subtracting twice the standard error from the point estimate (or twice the posterior standard deviation from the posterior mean, in the Bayesian mixture model).

The GEEPERs results suggest little difference, if any, between the effects in the two principal strata. Estimated effects compared to BAU or DragonBox were slightly lower for bottom-outers than for non-bottom-outers, and estimated effects compared to FH2T were higher for non-bottom-outers.
However, none of these differences is statistically significant, so opposite patterns, or no difference between principal effects at all, would also be consistent with the data.
In fact, 95\% confidence intervals for all principal effects include both positive and negative principal effects.

The other estimation strategies largely agreed with GEEPERs, and with each other. 
This may suggest that the estimators' various assumptions---additivity of treatment assignment effect for GEEPERs and the mixture model, normal residuals for the mixture model, and principal ignorability for PSW---were all innocuous and may have held approximately. Alternatively, violations of the different assumptions may have led to similar errors.
